Environmental and geographic variables are effective surrogates for genetic variation in conservation planning

### 2016-09-28 code/rmarkdown/article.Rmd
+ incorporated rich's comments on the ms

<!-- Introduction -->
Protected areas spearhead conservation efforts [@r440]. These places buffer species from anthropogenic impacts and provide places for species to persist. Since the resources available for conservation are limited, protected areas need to be sited in places that fulfill conservation objectives for minimal cost [@r255]. To achieve this, conservation planning exercises often generate plans for entire networks of protected areas (prioritisations) to allocate resources cost-effectively. By generating plans for multiple reserves simultaneously, protected areas can preserve broad-scale biodiversity processes.

Intra-specific genetic variation is important for long-term species persistence [@r165]. As a consequence, there has been increasing interest in designing prioritisations that preserve and cultivate this aspect of biodiversity [@r67; @r3; @r61; @r499]. Broadly speaking, although the strength of natural selection is a continuous force, genetic variation can be classified as either adaptive or neutral [reviewed in @r480]. Adaptive genetic variation is associated with loci that significantly affect fitness. Typically, such "adaptive" loci are detected due to anomalous patterns among individuals [eg. @r456; @r487; @r480; @r490], though phenotypic-genetic associations are more likely to identify functionally important genomic regions for individual species [eg. @r500]. By preserving existing patterns of adaptive variation, protected areas can ensure that populations with particularly beneficial adaptations are not lost [@r3; @r43]. In contrast, neutral genetic variation is associated with loci that do not significantly affect fitness. Neutral variation often reflects the evolutionary history of different populations. By preserving existing patterns of neutral variation, protected areas can safeguard against the adverse effects of low genetic diversity [eg. inbreeding depression; @r67]. Thus protected area networks should aim to preserve both adaptive and neutral patterns of genetic variation for species of conservation interest [@r67; @r3; @r43]. Despite this, genetic data are not often used to inform conservation planning exercises because such data require considerable investment to obtain and analyze [@r61; but see @r5 and @r157 for recent efforts]. Thus prioritisations may fail to sample species' intra-specific genetic variation, and ultimately fail to deliver long-term conservation outcomes.

Recently, the use of surrogate data has been proposed to generate prioritisations that adequately preserve intra-specific genetic variation--without needing to utilise genetic data to achieve this. Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been proposed to guide reserve selection as surrogates for adaptive genetic variation [eg. environmental conditions; @r202; @r247; @r2]. By preserving a more representative sample of environmental conditions across the species' ranges, prioritisations could be generated that preserve more adaptive variation. However, to our knowledge, the effectiveness of this approach remains unverified. On the other hand, neutral genetic variation arises from a reduction in gene flow between individuals. Over the last few years, the field of phylogeography has predominantly focused on describing neutral genetic diversity, and so the landscape factors that affect neutral variation are relatively well understood [@r501]. Potential surrogates for neutral variation have been based on variables that predict the levels of connectivity between planning units. For instance, the isolation by distance theory [IBD; @r259] predicts that populations further apart will have less connectivity, exchange less genetic material, and in turn have less genetic variation in common. Based on this theory, geographic distance-based surrogates have been found to improve prioritisations in an island network [@r200]. However this remains untested in spatially contiguous systems where connectivity is complicated by additional factors [eg. terrain or habitat type; @r474; @r497]--the very systems commonly featured in real-world conservation planning exercises [eg. @r197]. Before conservation planners can reliably use surrogates for genetic data in conservation planning, the effectiveness of these surrogates need to be verified. Otherwise, conservation planners may waste precious resources on protecting places that provide little benefit to biodiversity.

Here we investigated the effectiveness of environmental and geographic surrogates for adaptive and neutral genetic variation in the context of conservation planning. To achieve this goal, we used distribution and genomic data for 27 alpine plants collected by the international IntraBioDiv project in the European Alps [@r451]. We show that for most species the spatial patterns of adaptive and neutral genetic variation correlated with environmental and geographic variables. We also show that for most species the level of correlation is high enough to be operational for conservation planning. Finally, we demonstrate the using these surrogates results in prioritisations that secure more genetic variation.

@@

<!-- summary of methods -->
We detected putatively adaptive genetic variation in `r as.character(as.english(length(env.surrogacy.LST)))` of the 27 plant species. We used two outlier detection methods to detect loci under selection, and they returned relatively consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD). Of the species that were associated with adaptive genetic variation, only a small proportion of loci were classified as being adaptive (mean `r round(mean(spp.outlier.loci.props, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.loci.props, na.rm=TRUE),2)` SD). After identifying the loci showing strong signals of selection, we used non-metric multi-dimensional scaling (NMDS) analyses to identify the main gradients of adaptive (if detected) and neutral variation for each species. Generally, only a couple of continuous dimensions were needed to sufficiently describe their patterns of adaptive (all $K = `r unique(na.omit(adaptive.k))`$; Table S2) and neutral genetic variation (minimum $K = `r min(neutral.k)`$; mean $K = `r round(mean(neutral.k),2)`$; maximum $K = `r max(neutral.k)`$; Table S2). The resulting ordinations were used to construct an adaptive (if detected) and neutral genetic space for each species. The spatial distribution of these genetic spaces (Figures S29--S54) generally showed patterns of spatial auto-correlation (eg. _Cerastium uniflorum_, Figure S34; _Dryas octopetala_, Figures S36)--a good sign that environmental and geographic variables may prove effective surrogates.

@@

We verified that the spatial patterns in environmental and geographic variation correlated with the broad-scale spatial patterns in adaptive and neutral variation for most species (respectively). We constructed dissimilarity matrices between the planning units. For each species, we constructed matrices expressing the differences between planning units as a function of their environmental characteristics and geographic position, and the adaptive (if detected) and neutral genetic characteristics between individuals. We found the spatial patterns of environmental variation to be significantly correlated with those of adaptive genetic variation for `r as.character(as.english(sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)))` species of the `r as.character(as.english(sum(!is.na(correlation.model.summary.DF$Environmental.corr.P))))` species associated with adaptive variation (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models). We also found the geographic distance between planning units to significantly correlate with the spatial patterns of neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` species of the total `r sum(!is.na(correlation.model.summary.DF$Geographic.corr.P))` (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models). Thus, for most species, planning units that contained different environmental conditions contained individuals with different adaptive genetic characteristics, and planning units that were further apart tended to contain individuals with different neutral genetic characteristics. After verifying these patterns, our next step was to determine how effective these variables were as surrogates for conservation planning.

@@

<!-- surrogacy analysis -->
We found the environmental and geographic variables to be moderately effective predictors for genetic variation when assessing the performance of prioritisations (Figure 1). For each species, we randomly generated prioritisations, and calculated the proprotion of environmental, geographic, adaptive (if present), and neutral genetic variation they sampled. We found the relationship between the proportion of adaptive genetic variation and environmental variation secured in a given prioritisation to depend on the species for which the prioritisation was generated (species $\times$ proportion of environmental variation interaction term; ${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$). Similarly, we found the relationship between the proportion of neutral genetic variation and geographic variation secured in a prioritisation to depend on the species for which the prioritisation was generated (${\chi^2}_{`r round(geo.surrogacy.AOV$df[[4]],2)`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`). For all species, we found these relationship to be significantly different to zero (environmental: minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; geographic: minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3). After establishing that the surrogate variables predicted how well prioritisations secured genetic variation, our final step was to determine if prioritisations were more effective when generated using surrogate-based targets.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=6, fig.height=4.5, out.width='3.42in', out.height='2.5in', fig.cap='The relationship between surrogates and genetic variation secured in prioritizations. Panel (a) shows the relationship between the proportion of intra-specific adaptive genetic variation and environmental variation secured in a prioritization. Similarly, panel (b) shows the relationship between the proportion of intra-specific neutral variaiton and geograpgic variation secured in a prioritization. Panels (c) and (d) show the distribution of $R^2$ values for these models.'}
# make figure
p1 <- ggplot(aes(x=surrogate.held,y=genetic.held), data=env.surrogacy.DF) +
	xlab('Environmental variation\nsecured (%)') + 
	ylab('Adaptive genetic\nvariation secured (%)') +
	theme_classic() + theme(axis.line.x=element_line(color='black'),
		axis.line.y=element_line(color='black')) +
	coord_cartesian(xlim=c(0, 1))
for (i in seq_along(env.surrogacy.LST))
	p1 <- p1 + stat_smooth(method='glm', method.args=list(family='binomial'),
		data=filter(env.surrogacy.DF, Species==unique(Species)[i]),
		fullrange=TRUE, se=FALSE, colour='black')
p1 <- p1 + geom_text(x=-Inf, y=Inf, hjust=-0.25, vjust=1.5, label='(a)')

p2 <- ggplot(aes(x=surrogate.held,y=genetic.held), data=geo.surrogacy.DF) +
	xlab('Geographic variation\nsecured (%)') + 
	ylab('Neutral genetic\nvariation secured (%)') +
	theme_classic() + theme(axis.line.x=element_line(color='black'),
		axis.line.y=element_line(color='black', colour='transparent')) +
	coord_cartesian(xlim=c(0, 1))
for (i in seq_along(geo.surrogacy.LST))
	p2 <- p2 + stat_smooth(method='glm', method.args=list(family='binomial'),
		data=filter(geo.surrogacy.DF, Species==unique(Species)[i]),
		fullrange=TRUE, se=FALSE, colour='black')
p2 <- p2 + geom_text(x=-Inf, y=Inf, hjust=-0.25, vjust=1.5, label='(b)')

@@

<!-- scenario analysis -->
To determine if the addition of environmental and geographic targets could actually improve the effectiveness of prioritisations, we generated prioritisations according to three scenarios using different targets (Figure S58). We then computed the proportion of adaptive and neutral genetic variation secured in each prioritisation (Figure 3). We found no evidence to suggest that any interaction terms were significant (${\chi^2}_{`r scenario.ANOVA.1.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.1.DF$p.value[[2]])`$). This result indicates the trends observed in each of the main effects were consistent across all other levels in the main effects.

@@

First, the prioritisations tended to secure more adaptive (mean `r round(mean(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` % secured in all prioritisations $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)`) than neutral genetic variation (mean `r round(mean(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)` % secured in all prioritisations $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)`; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.3.DF$p.value[[2]])`). This result indicates that regardless of the targets used to generate a prioritisation, or the planning context under which the prioritisation was generated, prioritisations tend to secure more adaptive then neutral genetic variation.

Second, the proportion of genetic variation secured in a prioritisation was found to depend on the targets used to generate the prioritisation (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$p.value[[2]])`)$. Specifically, the amount-based prioritisations (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)` % overall genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)`) tended to secure less variation than the prioritisations generated using additional surrogate targets (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` % secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$) or genetic targets (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` % secured $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[2]])`$). No evidence was found to suggest that the prioritisations that included additional surrogate-based targets (`r raptr.params.LST[[MODE]]$scenario.analysis$surrogate.target*100` %) or genetic-based targets (`r raptr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %) secured a different proportion of genetic variation ($Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$). These results suggest that the addition of environmental and geographic surrogates can improve the proportion of intra-specific genetic variation secured in a prioritisation for most species.

Third, we found significant differences in the average proportion of genetic variation secured by prioritisations generated under different scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.2.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.2.DF$p.value[[2]])`). Specifically, both the multi-species (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)`) and multi-species with cost (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)`) were both found to secure more genetic variation than the single species scenario (mean `r round(mean(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held),2)`; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[2]])`$; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$ respectively). Neither of the multi-species scenarios were found to secure a different proportion of genetic variation ($Z_1 = `r round(scenario.GLHT.1.DF$statistic[[3]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[3]])`$). These results suggest that on average the multi-species prioritisations can secure more genetic variation than single-species prioritisations--regardless of targets used to generate them.

@@

<!--  introduction -->
We aimed to investigate the effectiveness of broad-scale environmental and geographic variables as surrogates for adaptive and neutral genetic variation in conservation planning. Our study is the first to address this aim using a diverse set of species and actually measuring the proportion of genetic variation secured in prioritisations. Overall, we found that the spatial patterns of environmental and geographic variables correlate with those of adaptive and neutral patterns for most species. Furthermore, we found that the addition of environmental and geographic targets to the reserve selection process resulted in prioritisations that secured a greater proportion of adaptive and neutral genetic variation for most species.

>>>>

<!-- Introduction -->
Protected areas spearhead conservation efforts [@r440]. They buffer species from anthropogenic impacts, providing places for species to persist. Since the resources available for conservation are limited, protected areas need to be sited in places that fulfill conservation objectives for minimal cost [@r255]. To achieve this, conservation planning exercises often generate plans for entire networks of protected areas (prioritisations) to allocate resources cost-effectively and preserve broad-scale biodiversity processes. Intra-specific genetic variation is fundamental to long-term species persistence [@r165], and as a consequence, there has been increasing interest in designing prioritisations that capture and cultivate this aspect of biodiversity [@r67; @r3; @r61; @r499].

Broadly speaking, though the strength of natural selection is a continuous force, genetic variation can be classified as either adaptive or neutral [reviewed in @r480]. Adaptive genetic variation is associated with loci that significantly affect fitness. Typically, such "adaptive" loci are detected due to anomalous patterns among individuals [eg. @r456; @r487; @r480; @r490], though phenotypic-genetic associations are more likely to identify functionally important genomic regions for individual species [eg. @r500]. If they represent the full range of adaptive variation, protected areas can help ensure that populations with particularly beneficial adaptations are not lost [@r3; @r43]. In contrast, neutral genetic variation is associated with loci that do not significantly affect fitness, but instead reflects the evolutionary history of different populations. By representing the full range of neutral variation, protected areas can safeguard against the adverse effects of low genetic diversity [@rXXX]. Thus optimally sited protected area networks would aim to represent both adaptive and neutral patterns of genetic variation for species of conservation interest [@r67; @r3; @r43].

Recently, the use of surrogate data has been proposed to generate prioritisations that adequately preserve intra-specific genetic variation--without needing to directly utilise genetic data to achieve this. Since adaptive genetic variation is ultimately driven by extrinsic selection pressures, environmental variables have been proposed to guide reserve selection as surrogates for adaptive genetic variation [eg. environmental conditions; @r202; @r247; @r2]. By preserving a more representative sample of environmental conditions across the species' ranges, prioritisations could be generated that also capture more adaptive variation. However, the effectiveness of environmental surrogates in representing adaptive genetic variation remains unverified. Neutral genetic variation arises from a reduction in gene flow between individuals. Over the last few years, the field of phylogeography has predominantly focused on describing neutral genetic diversity, and so the landscape factors that affect neutral variation are relatively well understood [@r501], and potential surrogates for neutral variation have been based on variables that predict the levels of connectivity between planning units. For instance, isolation by distance [IBD; @r259] predicts that populations further apart will be more weakly connected, exchange less genetic material, and in turn have less genetic variation in common. Based on this idea, geographic surrogates based on distance have been found to improve conservation prioritisations in an island network [@r200]. However this remains untested in spatially contiguous systems where connectivity is complicated by additional factors [eg. terrain or habitat type; @r474; @r497]--the typical situations in most conservation planning exercises [eg. @r197].

Here we investigate the effectiveness of environmental and geographic surrogates for capturing adaptive and neutral genetic variation respectively in the context of conservation planning. We used distribution and genomic data for 27 alpine plants collected by the international IntraBioDiv project in the European Alps [@r451]. We show that for most species the spatial patterns of adaptive and neutral genetic variation correlated with environmental and geographic variables. We also show that for most species the level of association is strong enough to be operational for conservation planning. Finally, we demonstrate the using these surrogates results in prioritisations that secure more genetic variation than expected by chance.

@@

<!-- summary of methods -->
We detected putatively adaptive genetic variation in `r as.character(as.english(length(env.surrogacy.LST)))` of the 27 plant species. We used two outlier detection methods to detect loci under selection, and they returned reasonably consistent results (mean `r round(mean(spp.outlier.methods.congruence, na.rm=TRUE),2)` % loci per species assigned the same classification $\pm$ `r round(sd(spp.outlier.methods.congruence, na.rm=TRUE),2)` SD). Of the species that were associated with adaptive genetic variation, only a small proportion of loci were classified as being adaptive (mean `r round(mean(spp.outlier.loci.props, na.rm=TRUE),2)` % loci per species $\pm$ `r round(sd(spp.outlier.loci.props, na.rm=TRUE),2)` SD). After identifying the loci showing strong signals of selection, we used non-metric multi-dimensional scaling (NMDS) analyses to identify the main gradients of adaptive (if detected) and neutral variation for each species. Generally, only a small number of continuous dimensions was needed to sufficiently describe their patterns of adaptive (all $K = `r unique(na.omit(adaptive.k))`$; Table S2) and neutral genetic variation (minimum $K = `r min(neutral.k)`$; mean $K = `r round(mean(neutral.k),2)`$; maximum $K = `r max(neutral.k)`$; Table S2). The resulting ordinations were used to construct an adaptive (if detected) and neutral genetic space for each species. The spatial distribution of these genetic spaces (Figures S29--S54) generally showed substiantial spatial auto-correlation (eg. _Cerastium uniflorum_, Figure S34; _Dryas octopetala_, Figures S36), suggesting that the environmental and geographic variables may prove to be effective surrogates.

@@

We verified that the spatial patterns in environmental and geographic variation correlated with the broad-scale spatial patterns in adaptive and neutral variation for most species (respectively). We constructed dissimilarity matrices between the planning units. For each species, we constructed matrices expressing the differences between planning units as a function of their environmental characteristics and geographic position, and the adaptive (if detected) and neutral genetic characteristics between individuals. The spatial patterns of environmental variation was significantly correlated with that of adaptive genetic variation for `r as.character(as.english(sum(correlation.model.summary.DF$Environmental.corr.P<0.05,na.rm=TRUE)))` of the `r as.character(as.english(sum(!is.na(correlation.model.summary.DF$Environmental.corr.P))))` species associated with adaptive variation (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Environmental.corr.P<0.05)$Environmental.Marginal.R2),2)` SD for significant models). Geographic distance between planning units also correlated significantly with the spatial pattern of neutral genetic variation among planning units for `r sum(correlation.model.summary.DF$Geographic.corr.P<0.05,na.rm=TRUE)` of the `r sum(!is.na(correlation.model.summary.DF$Geographic.corr.P))` species (_P_ < 0.05; mean `r round(mean(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` marginal $R^2$ $\pm$ `r round(sd(filter(correlation.model.summary.DF, Geographic.corr.P<0.05)$Geographic.Marginal.R2),2)` SD for significant models). Thus, for most species, planning units that contained different environmental conditions contained individuals with different adaptive genetic characteristics, and planning units that were further apart tended to contain individuals with different neutral genetic characteristics. After verifying these patterns, our next step was to determine whether these variables are effective surrogates for conservation planning.

@@

<!-- surrogacy analysis -->
For each species, we generated a suite of `r raptr.params.LST[[MODE]]$surrogacy.analysis$random.replicates*2` random prioritisations, and calcuated the proportion of environmental variation and adaptive (if present) genetic variation they sampled. We then repeated this process, and calculated the proportion of geographic variation and neutral genetic variation they sampled. Environmental and geographic variables were moderately effective predictors for genetic variation for the genetic variation represented by the randomly generated prioritisations (Figure 1). The relationship between the proportion of genetic variation and surrogate variation secured in the prioritisations varied among species for environmental surrogates of adaptive genetic variation (species $\times$ proportion of environmental variation interaction term; ${\chi^2}_{`r env.surrogacy.AOV$df[[4]]`} = {`r round(env.surrogacy.AOV$Resid..Dev[[4]],2)`}; P `r pvalString(env.surrogacy.AOV$Pr..Chi.[[4]])`$), and geographic surrogates of neutral variation (${\chi^2}_{`r round(geo.surrogacy.AOV$df[[4]],2)`} = `r round(geo.surrogacy.AOV$Resid..Dev[[4]],2)`$; $P$ `r pvalString(geo.surrogacy.AOV$Pr..Chi.[[4]])`). _Post-hoc_ analyses showed that the relationships for all species were significantly different to zero (environmental: minimum \textit{z} = `r round(max(env.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; geographic: minimum \textit{z} = `r round(min(geo.surrogate.spp.DF$statistic),2)`, maximum $P `r pvalString(max(env.surrogate.spp.DF$p.value))`$; Table S3). After establishing that the surrogate variables predicted how well prioritisations secured genetic variation, our final step was to determine if prioritisations were more effective when generated using surrogate-based targets.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=6, fig.height=4.5, out.width='3.42in', out.height='2.5in', fig.cap='Adaptive (a) and neutral (b) genetic variation represented in a suite of randomly generated prioritisations as a function of the environmental and geographic surrogates they also secured (respectively). Panels (c) and (d) show the distribution of $R^2$ values for models using each surrogate to explain genetic variation.'}
# make figure
p1 <- ggplot(aes(x=surrogate.held,y=genetic.held), data=env.surrogacy.DF) +
	xlab('Environmental variation\nsecured (%)') + 
	ylab('Adaptive genetic\nvariation secured (%)') +
	theme_classic() + theme(axis.line.x=element_line(color='black'),
		axis.line.y=element_line(color='black')) +
	coord_cartesian(xlim=c(0, 1)) +
	scale_x_continuous(limits=c(0,1), labels=function(x) {x*100}) +
	scale_y_continuous(limits=c(0,1), labels=function(x) {x*100})
for (i in seq_along(env.surrogacy.LST))
	p1 <- p1 + stat_smooth(method='glm', method.args=list(family='binomial'),
		data=filter(env.surrogacy.DF, Species==unique(Species)[i]),
		fullrange=TRUE, se=FALSE, colour='black')
p1 <- p1 + geom_text(x=-Inf, y=Inf, hjust=-0.25, vjust=1.5, label='(a)')

p2 <- ggplot(aes(x=surrogate.held,y=genetic.held), data=geo.surrogacy.DF) +
	xlab('Geographic variation\nsecured (%)') + 
	ylab('Neutral genetic\nvariation secured (%)') +
	theme_classic() + theme(axis.line.x=element_line(color='black'),
		axis.line.y=element_line(color='black', colour='transparent')) +
	coord_cartesian(xlim=c(0, 1)) +
	scale_x_continuous(limits=c(0,1), labels=function(x) {x*100}) +
	scale_y_continuous(limits=c(0,1), labels=function(x) {x*100})
for (i in seq_along(geo.surrogacy.LST))
	p2 <- p2 + stat_smooth(method='glm', method.args=list(family='binomial'),
		data=filter(geo.surrogacy.DF, Species==unique(Species)[i]),
		fullrange=TRUE, se=FALSE, colour='black')
p2 <- p2 + geom_text(x=-Inf, y=Inf, hjust=-0.25, vjust=1.5, label='(b)')

@@

<!-- scenario analysis -->
To determine whether using environmental and geographic targets could improve the effectiveness of prioritisations in representing genetic variation, we generated prioritisations using (i) "amount targets" reflecting the traditional approach of representing a certain proportion of each species' goegraphic distribution, (ii) "amount and surrogate targets" where we also targeted representation of the environmental and geographic surrogates as well as a certain proportion of each species' geographic distribution, and (iii) "amount and genetic targets" where we targeted representation of the directly measured genetic variation as well as a certain proportion of each species' geographic distribution (Figure 3). To determine if these patterns were robust, for each of the three combinations of the targets, we generated prioritsations under four scenarios: (i) single-species with equal costs, (ii) single species with opportunity costs, (iii) multi-species with equal costs, and (iv) multi-species with opportunity costs.

@@

The proportion of genetic variation secured in a prioritisation depended on the targets used to generate the prioritisation (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$p.value[[2]])`)$. Most notably, prioritisations based on the surrogate variables represented significantly more genetic variation than amount-based prioritisations (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & surrogate\ntargets')$genetic.held*100),2)` overall genetic variation secured versus `r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount targets')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$), and were not distinguishable from prioritisations based on measured genetic data (`r round(mean(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Prioritisation=='Amount & genetic\ntargets')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$). These results indicate that environmental and geographic surrogates in this case far outperform traditional amount-based conservation planning, and perform as well as a conservation plan based directly on measured genetic data.

Overall, the prioritisations tended to secure more adaptive then neutral genetic variation (adaptive: mean `r round(mean(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` % $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Adaptive variation')$genetic.held*100),2)` secured across all prioritisations, neutral: mean `r round(mean(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)` $\pm$ `r round(sd(filter(scenario.sub.DF,Metric=='Neutral variation')$genetic.held*100),2)`; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.3.DF$p.value[[2]])`). Thus, regardless of the targets used to generate a prioritisation, or the planning context under which the prioritisation was generated, prioritisations tend to secure more adaptive than neutral genetic genetic variation.

The average proportion of genetic variation secured by prioritisations varied under different scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.2.DF$statistic[[2]],2)`$; _P_ `r pvalString(scenario.ANOVA.2.DF$p.value[[2]])`). Both the multi-species (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(equal costs)')$genetic.held*100),2)`) and multi-species with cost (mean `r round(mean(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Multi-species\n(opportunity costs)')$genetic.held*100),2)`) were both found to secure more genetic variation than the single species scenario (mean `r round(mean(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held*100),2)` % genetic variation secured $\pm$ `r round(sd(filter(scenario.sub.DF,Context=='Single-species\n(equal costs)')$genetic.held*100),2)`; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[2]])`$; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$ respectively). These results suggest that on average the multi-species prioritisations can secure more genetic variation than single-species prioritisations--regardless of targets used to generate them.

@@

<!--  introduction -->
We aimed to investigate the effectiveness of broad-scale environmental and geographic variables as surrogates for adaptive and neutral genetic variation in conservation planning. Our study is the first to address this aim using a diverse set of species and calibrated against field measurements of genetic variation. The spatial patterns of environmental and geographic surrogates was strongly correlated with the patterns of adaptive and neutral genetic variation for most species, and environmental and geographic targets resulted in prioritisations that secured a greater proportion of adaptive and neutral genetic variation than traditional conservation planning for most species. These environmental and geographic surrogates are based on freely available data, and could be calculated for any study region across the world.
